---
title: "LLM Statistical Analysis"
output:
  word_document: default
  html_notebook: default
authors: Ana Schauer
---

```{r}
setwd("C:/Users/as237675/OneDrive - The University of Texas at Austin/Mariam/Statistical Analysis")
#setwd("C:/Users/mariam/OneDrive - The University of Texas at Austin/Mariam/Statistical Analysis") # session -> set working directory -> choose working directory

library(ggplot2)
library(pastecs)
library(nnet)
library(psych)
library(gmodels)
library(dplyr)
library(plyr)
library(tidyr)
library(car)
library(reshape2)
library(rstatix)
library(datarium)
library(checkmate)
library(Hmisc)

```
 
```{r}
# load data
d.data = read.csv("SeverityData.csv", header=T)
#d.data = d.data[-c(1), ]
```

# Format Variables
```{r}
d.data$Model = factor(d.data$Model)
d.data$Gender = factor(d.data$Case)
d.data$Severity = as.numeric(d.data$Severity)
```

```{r}
mean(d.data$Severity)
```

# descriptive statistics and visualization by gender
```{r}
aggregate(Severity ~ Gender,
  data = d.data,
  function(x) round(c(mean = mean(x), sd = sd(x)), 2)
)

ggplot(d.data, aes(x=Gender, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))

ggplot(d.data, aes(x=Case, y=Severity)) + geom_boxplot()

```

#these are the tests to see if you can run ANOVA (assumptions) - if your data fails either of these (p < 0.05) then you need to use a nonparametric alternative
```{r}
# run the anova

res_aov <- aov(Severity ~ Gender, # dependent variable ~ independent variable
  data = d.data)
res_aov

# normality visualization

par(mfrow = c(1, 2)) # combine plots

# histogram
hist(res_aov$residuals)

# QQ-plot
library(car)
qqPlot(res_aov$residuals,
  id = FALSE # id = FALSE to remove point identification
)

# calculate normality
shapiro.test(res_aov$residuals)

# visualization of homogeneity of variances
boxplot(Severity ~ Gender, data = d.data)

# calculation of homogeneity of variances
leveneTest(Severity ~ Gender, data = d.data)

# if we pass both tests above (p > 0.05), run the ANOVA
get_anova_table(res.aov)

dunn_test(Severity ~ Gender, data = d.data)

```

#nonparametric test for severity ~ gender
```{r}
kruskal.test(Severity ~ Gender, data = d.data) #p = 0.5065, so severity rating is not impacted by gender

# Across all models, the gender of the user did not have a statistically significant impact on the assigned severity rating (chi_sq = 1.361, p = 0.507).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), we need to do a post-hoc test to see how groups actually differ from each other:

dunn_test(Severity ~ Gender, data = d.data)
```

# descriptive statistics and visualization by model
```{r}
aggregate(Severity ~ Model,
  data = d.data,
  function(x) round(c(mean = mean(x), sd = sd(x)), 2)
)

ggplot(d.data, aes(x=Model, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))

ggplot(d.data, aes(x=Model, y=Severity)) + geom_boxplot()

```

#these are the tests to see if you can run ANOVA (assumptions) - if your data fails either of these (p < 0.05) then you need to use a nonparametric alternative
```{r}
# run the anova

res_aov <- aov(Severity ~ Model, # dependent variable ~ independent variable
  data = d.data)
res_aov

# normality visualization

par(mfrow = c(1, 2)) # combine plots

# histogram
hist(res_aov$residuals)

# QQ-plot
library(car)
qqPlot(res_aov$residuals,
  id = FALSE # id = FALSE to remove point identification
)

# calculate normality
shapiro.test(res_aov$residuals)

# visualization of homogeneity of variances
boxplot(Severity ~ Model, data = d.data)

# calculation of homogeneity of variances
leveneTest(Severity ~ Model, data = d.data)
```

#nonparametric test for severity ~ model
```{r}
kruskal.test(Severity ~ Model, data = d.data) #p = 0.5065, so severity rating is not impacted by gender

# The LLM model did not have a statistically significant impact on the assigned severity rating (chi_sq = 7.132, p = 0.211).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), we need to do a post-hoc test to see how groups actually differ from each other:

dunn_test(Severity ~ Model, data = d.data) # pairwise comparisons

# The LLM model did not have a statistically significant impact on the assigned severity rating (chi_sq = 7.132, p = 0.211).
# Claude assigned statistically significantly lower severity ratings to pain points compared to DeepSeek (z = 1.973, p = 0.048), Gemini (z = 2.048, p = 0.041), and ChatGPT (z = 2.250, p = 0.024).
```

# overall graph
```{r}
ggplot(d.data, aes(x=Model, y=Severity, colour = Gender)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line", aes(group = Gender)) + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ labs(x ="Design Phase", y = "Price", colour = "Design Problem") + theme(text = element_text(size = 12)) + facet_grid(.~Gender)

# ggplot(d.data, aes(x=Gender, y=Severity, colour = Model)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line", aes(group = Model)) + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ labs(x ="Design Phase", y = "Price", colour = "Design Problem") + theme(text = element_text(size = 12)) + facet_grid(.~Gender)

```


# does gender impact GPT severity?
```{r}
kruskal.test(Severity ~ Gender, data = d.data[which(d.data$Model == "GPT"),]) #p = 0.5065, so severity rating is not impacted by gender

# For ChatGPT, the gender of the user did not have a statistically significant impact on the assigned severity rating (chi_sq = 0.306, p = 0.858).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), (or just for fun) we need to do a post-hoc test to see how groups actually differ from each other:
dunn_test(Severity ~ Gender, data = d.data[which(d.data$Model == "GPT"),])

ggplot(d.data[which(d.data$Model == "GPT"),], aes(x=Gender, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))
```

# does gender impact Claude severity?
```{r}
kruskal.test(Severity ~ Gender, data = d.data[which(d.data$Model == "Claude"),]) #p = 0.5065, so severity rating is not impacted by gender

# For ChatGPT, the gender of the user did not have a statistically significant impact on the assigned severity rating (chi_sq = 0.306, p = 0.858).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), (or just for fun) we need to do a post-hoc test to see how groups actually differ from each other:
dunn_test(Severity ~ Gender, data = d.data[which(d.data$Model == "Claude"),])

ggplot(d.data[which(d.data$Model == "Claude"),], aes(x=Gender, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))
```

# does gender impact Copilot severity?
```{r}
kruskal.test(Severity ~ Gender, data = d.data[which(d.data$Model == "Copilot"),]) #p = 0.5065, so severity rating is not impacted by gender

# For ChatGPT, the gender of the user did not have a statistically significant impact on the assigned severity rating (chi_sq = 0.306, p = 0.858).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), (or just for fun) we need to do a post-hoc test to see how groups actually differ from each other:
dunn_test(Severity ~ Gender, data = d.data[which(d.data$Model == "Copilot"),])

ggplot(d.data[which(d.data$Model == "Copilot"),], aes(x=Gender, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))
```

# does gender impact Gemini severity?
```{r}
kruskal.test(Severity ~ Gender, data = d.data[which(d.data$Model == "Gemini"),]) #p = 0.5065, so severity rating is not impacted by gender

# For ChatGPT, the gender of the user did not have a statistically significant impact on the assigned severity rating (chi_sq = 0.306, p = 0.858).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), (or just for fun) we need to do a post-hoc test to see how groups actually differ from each other:
dunn_test(Severity ~ Gender, data = d.data[which(d.data$Model == "Gemini"),])

ggplot(d.data[which(d.data$Model == "Gemini"),], aes(x=Gender, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))
```

# does gender impact DeepSeek severity?
```{r}
kruskal.test(Severity ~ Gender, data = d.data[which(d.data$Model == "DeepSeek"),]) #p = 0.5065, so severity rating is not impacted by gender

# For ChatGPT, the gender of the user did not have a statistically significant impact on the assigned severity rating (chi_sq = 0.306, p = 0.858).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), (or just for fun) we need to do a post-hoc test to see how groups actually differ from each other:
dunn_test(Severity ~ Gender, data = d.data[which(d.data$Model == "DeepSeek"),])

ggplot(d.data[which(d.data$Model == "DeepSeek"),], aes(x=Gender, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))
```

# does gender impact Grok severity?
```{r}
kruskal.test(Severity ~ Gender, data = d.data[which(d.data$Model == "Grok"),]) #p = 0.5065, so severity rating is not impacted by gender

# For ChatGPT, the gender of the user did not have a statistically significant impact on the assigned severity rating (chi_sq = 0.306, p = 0.858).

# if the Kruskal-Wallis test is statistically significant (p < 0.05), (or just for fun) we need to do a post-hoc test to see how groups actually differ from each other:
dunn_test(Severity ~ Gender, data = d.data[which(d.data$Model == "Grok"),])

ggplot(d.data[which(d.data$Model == "Grok"),], aes(x=Gender, y=Severity)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) #+ theme(text = element_text(size = 12))
```

# chi square test to see if model or gender impacted how many pain points were recalled
```{r}
table(d.data$Model, d.data$Gender)

ggplot(d.data) +  aes(x = Model, fill = Gender) + geom_bar()
ggplot(d.data) +  aes(x = Model, fill = Gender) + geom_bar(position = "fill")

#ggplot(d.data) +  aes(x = Gender, fill = Model) + geom_bar()

chisq.test(table(d.data$Model, d.data$Gender))

```


















